\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-10 concepts.tex
finished cleanup of "EUROPA plan representation section". Still need figures and a few minor notes remain.

\item[\textbf{Variables}] Values that need to be represented to
  describe the problem domain and over which we may want to specify
  constraints. In the Shopping Agent example, the times at which the
  agent needs to leave or be back, or executes a purchase, are all
  instances where variables representing time would be used. 
  As explained in the CP section, a variable can take values from a domain. In \eu, domains are defined over a specific data type, the primitive data types supported are:
  \begin{enumerate}
    \item \textit{String}: sequences of characters, like "Red", "Spirit", etc 
    \item \textit{Boolean}: {True,False}
    \item \textit{Numeric}: integers and floats
    \item \textit{Object}: reference to an object instance
  \end {enumerate}
  Two types of variable Domains can be defined over a data type: 
  \begin{enumerate}
    \item \textit{Enumerated}: Finite sets over any data type, specified explicitly, for instance: ["Red","Yellow","Green"], [1,18,32], etc
    \item \textit{Interval}: Only for Numeric data types, they are specified by a [LowerBound,UpperBound] pair, for instance: [1,10], [5.0,100.0], etc.
  \end {enumerate}
  \eu's ability to represent and reason over numeric intervals is a very useful characteristic that allow users to create flexible plans. \comment{say more about the benefits of flexibility?}

\item[\textbf{Objects}] The items we wish to describe and refer to in
  a domain are considered Objects. As in the case with object-oriented
  analysis and design, one can seek out the nouns in any domain
  description to find likely objects. In the Shopping Agent example,
  we night consider the Agent itself, Products and Locations all to be
  objects. Objects have state and behavior. For example, a Shopping
  Agent can have:

@@

  \item State: it's location, a bag that contains the products it has
    already purchased (the bag can in turn be another object), etc.

@@

  Similarly objects that have similar state and behavior can be
  described generically in terms of object types or classes. \eu
  allows the definition of object classes in the same way that it is
  done in popular object-oriented programming languages. However, to
  describe state and behavior for the purposes of planning we need to
  build on the formalism of first order logic as explained below.

@@

\comment {resume cleanup here}

\item[\textbf{Rules}] In order for a plan to be valid, it must comply with all rules and
regulations pertinent to the application domain in question. Rules
govern the internal and external relationships of a token. For
example, consider a parameterized predicate describing a transition
from one location to another. Let the parameters be from and to. The
parameters are instantiated on a token as variables whose domain of
values is the set of all locations in a given problem. A rule
governing an internal relationship among token variables might
stipulate that a transition must involve a change in location. This
can be easily expressed as a constraint on the definition of a
predicate of the form: from != to. It is reasonable to further
stipulate that one must be Located somewhere before a transition can
occur, and one must end up Located somewhere when completed. This is
an example of a rule governing an external relationship among
tokens. It specifies a requirement that tokens of the predicate
Located precede and succeed tokens of the predicate Moving.

Figure 2: Internal and external relationships from rules on Moving
Figure 2 illustrates the entities and relations involved in specifying
such a rule on a Moving predicate. The token on which the rule applies
is referred to as the master. Each Located token required by the
master is referred to as a slave. All variables are indicated by name
and their domains are expressed as intervals in the case of temporal
variables and as enumerations for the remainder. Application of a rule
on a token can thus cause slave tokens, variables, and constraints to
be introduced.

The capability of a domain rule to cause a slave token to be created
is a key vehicle through which planning occurs. Semantically, this
rule imposes a requirement for supporting tokens to be in the plan in
order for the master to be valid.

@@

\comment{instead of summary, tie every concept to the shopping example?}

This section has presented the main elements of \eu's planning representation:

\begin{enumerate}
    \item Temporally scoped predicates and actions (a.k.a. Tokens) to represent
    states and actions in time. Note that tokens do not discriminate between state and action.

    \item Token States which are the basis of planning operations on a partial plan.

    \item Constraints to describe relationships among tokens. This provides an
    expressive method of describing interactions among tokens in the context of Temporal Planning.

    \item Timelines as a concise abstraction to express the evolution of state
    and behavior for a system component. It provides semantics of mutual
    exclusion sequencing in time. Other core abstractions are available in
    \eu for handling metric resources.

    \item Domain rules to describe internal and external relationships on and
     between tokens respectively. Rules are applied to active tokens (master tokens) 
     referred to as masters and typically produce inactive
     tokens (slaves) which are a key vehicle through which the planning
     process evolves. The rule structure in \eu differs from the more
     restrictive commitment to preconditions and effects in classical
     planning which prohibits durative actions and disembodied effects
     (i.e. effects which may not occur until some temporal distance after
     the end of the action).
\end{enumerate}

%Each of the token variables, including the parameter variables, has a
%domain of values assigned to it.  The variables may also participate
%in constraints that specify which value combinations are valid.  To
%allow the specification of multiple values, e.g, to express a range of
%possible start times, variables are used to specify parameter, start
%and end time values for a token.  As a result, a token $T$ is a tuple
%$\langle v, P, s, e \rangle$, where $v$ is a variable denoting a state
%variable, $P$ is the name of a procedure and $s$ and $e$ (satisfying
%$s\leq e$) are variables encoded as flexible token start and end time
%points. Consequently, partial plans represent a \textit{range of
%  possible solutions} rather than an explicit definition of a single
%trajectory of state.  Fig. \ref{fig:flex-timelines} shows an example
%of a part of such plans with tokens representing constraints on
%flexible temporal intervals. Systematic computation is reflected by
%the use of the \texttt{Allen Algebra} \cite{allen84} temporal
%relations shown in Fig. \ref{fig:allen-algebra}. It is the use of such
%interval arithmetic in the \eu domain model that allows
%encoding of temporal constraint relationships between tokens on the
%same or other concurrent timelines. Details of propagation algorithms
%or of interval logic are beyond the scope of this paper.
%
%\begin{figure}
%\centering
%\includegraphics[scale=0.3]{figs/FSM-transition.pdf}
%\caption{\small Timeline evolution can be described by simple finite
%  state transitions where temporal relationships describe the arc
%  transition and temporal extant of state are token
%  durations. Constraints here are for \emph{meets} or its equivalent
%  inverse \emph{met\_by}.}
%\label{fig:FSM}
%\vskip-0.3cm
%\end{figure}
%
%In describing the evolution of state via timelines, transitions within
%a timeline can be described by simple finite state
%machines. Fig. \ref{fig:FSM} for example describes a timeline
%evolution for an abstract timeline entity describing navigation. Arc
%transitions represent temporal constraints and token duration are
%represented by temporal extant of state. The entire vehicle can then
%be modeled as a number of state transitions with concurrency for each
%of the modeled state variables defined a priori. However this simple
%view is often complicated when temporal constraints between state
%variables are articulated in our domain models. If there is a
%criticism of this approach, modeling then, requires a proper balance
%between abstraction, state variable description and concurrency
%described by the inter-connectedness between such finite state
%machines.

%For example, consider the tire-world domain. The tire was located in
%the trunk with the predicate tireLocated(Trunk) and located on the
%ground with the predicate tireLocated(Ground). Clearly, the same tire
%cannot be in both places at once. A Timeline provides a simple method
%for aggregating the statements about a tire such that they are
%mutually exclusive.
%
%Figure 1: A Timeline for a Tire
%
%Figure 1 illustrates how a Timeline can be used to describe the
%whereabouts of a tire. The tire is an object in the tire-world
%represented as a timeline. It has predicates associated with it which
%can describe its states and actions over time. The predicate names
%need not contain the tire prefix; this is now implicit since the
%tokens are assigned to a specific tire instance (i.e. an instance of a
%Timeline). While Timelines are a useful element of the \eu planning
%paradigm, they are not essential. A more general notion of a system
%Object can be used which does not impose restrictions of
%mutual-exclusion or non-zero duration. This can be important in
%supporting partial-order planning.
%
%For this example, the state of the tire is specified using 3
%contiguous tokens. The end and start time-points are related by an
%equality constraint. A Moving predicate has been introduced to cover
%the transition from one location to another. It takes 2 location
%arguments. Notice that precedence relationships exist between tokens
%such that they cannot overlap but the start and end times may remain
%flexible. To illustrate this, sample times are included. Assume a
%total time range of interest between 0 and 1000. In addition, tokens
%on Timelines have a minimum duration of 1. As a result the values
%shown are the most flexible possible values for the start and end of
%each token. There are a number of advantages of allowing this
%flexibility. First, the basic structure of the plan can be developed
%without over-committing to specific times. If it is not necessary for
%the tire to be on the ground at time-step 4, then a planner should not
%be forced to specify it. Such an approach permits a least-commitment
%approach to planning. Second, in many domains it is simply impossible
%to know in planning exactly how long an activity might take. For
%example, a common activity of driving a car from one point to another
%on a road can take varying amounts of time depending on traffic and
%traffic lights. In such circumstances it is more practical to express
%upper and lower bounds on durations which naturally lead to intervals
%for start and end times. In such domains, flexibility in planning aids
%robustness in execution.

\begin{figure*}
\centering 
\subfloat[]{\label{fig:plan-evolve1}\includegraphics[width=0.3\textwidth]{figs/Plan-evolve-1.pdf}} 
\subfloat[]{\label{fig:plan-evolve2}\includegraphics[width=0.3\textwidth]{figs/Plan-evolve-2.pdf}} 
\subfloat[]{\label{fig:plan-evolve3}\includegraphics[width=0.3\textwidth]{figs/Plan-evolve-3.pdf}} 
\caption{\small An illustrative plan synthesis example with concurrent
  timelines. Each timeline is an instantiation of a subsystem tracked
  by the planner. Tokens describe the instantiated state of subsystem
  at a particular time instance and enforce temporal and parametric
  constraints. Abstract goals are decomoposed into successively less
  abstract tokens and instantiated in co-temporal timelines using
  \texttt{Allen Algebra} relations. All tokens represent flexible
  start/end times. \ref{fig:plan-evolve1} shows an initial state
  evolving into \ref{fig:plan-evolve2} and \ref{fig:plan-evolve3}.}
  \label{fig:Plan-evolve}
  \vskip-5pt
\end{figure*}

Fig. \ref{fig:Plan-evolve} shows an illustrative example. We show four
timelines which track \texttt{Goal, Path, Navigation} and
\texttt{Command} state over time. Tokens on the goal timeline indicate
a survey within a bounded volume and a min/max depth envelope. As the
Volume Survey sub-goals on the \texttt{Path} timeline, the initial
token on that timeline gets \emph{squeezed} to have dependencies of
the goal token instantiated. Initially this dependency are the tokens
\texttt{Go(x1,y1)} and \texttt{Go(xn,yn)} indicative of the area of
coverage. The first of these in turn generates a sub-goal on the same
timeline to go to the next intermediate waypoint, which is illustrated
in Fig. \ref{fig:plan-evolve2}. Subsequent sub-goals (not shown in the
figure) ultimately generate all the tokens to completion for this
timeline. Fig. \ref {fig:plan-evolve3} shows a snapshot further along
in plan generation and shows additional subgoals on the \texttt{Path,
  Navigation} and \texttt{Command} timelines. Each token can thus
trace its causality when instantiated in the plan. The domain model
provides source of these temporal and parametric dependencies in the
form of rules.

\begin{minipage}[c]{\textwidth}
\vspace{+0.5cm}
% \framebox[\textwidth][t]{
Volume\_Survey(x$_1$,y$_1$,x$_n$,y$_n$,Min\_Depth, Max\_Depth)\\
$\Rightarrow$ \\
met\_by Go(x$_1$,y$_1$,Min\_Depth); \\
starts [50,0] Go(x$_2$,y$_2$,Min\_Depth,Max\_Depth); \\
\ldots{} \\
ends [50,0] Go(x$_{n-1}$,y$_{n-1}$, Min\_Depth,Max\_Depth); \\
meets Go(x$_n$,y$_n$,Min\_Depth, Max\_Depth); 
\end{minipage}

>>>>

\item[\textbf{Variables}] Values that need to be represented to
  describe the problem domain and over which we may want to specify
  constraints. In the Shopping Agent example, the times at which the
  agent needs to leave or be back, or executes a purchase, are all
  instances where variables representing time would be used. 
  As explained in the CP section, a variable can take values from a domain. In \eu, domains are defined over a specific data type, the primitive data types supported are:
  \begin{enumerate}
    \item \textit{String}: sequences of characters, like "Red", "Spirit", etc 
    \item \textit{Boolean}: {True,False}
    \item \textit{Numeric}: integers and floats
    \item \textit{Object}: reference to an object instance (see below for a description of objects in \eu)
  \end {enumerate}
  Two types of variable Domains can be defined over a data type: 
  \begin{enumerate}
    \item \textit{Enumerated}: Finite sets over any data type, specified explicitly, for instance: ["Red","Yellow","Green"], [1,18,32], etc
    \item \textit{Interval}: Only for Numeric data types, they are specified by a [LowerBound,UpperBound] pair, for instance: [1,10], [5.0,100.0], etc.
  \end {enumerate}
  \eu's ability to represent and reason over numeric intervals is a very useful characteristic that allow users to create flexible plans. \comment{say more about the benefits of flexibility?}

\item[\textbf{Constraints}] valid plans in most problem domains have to satisfy a number of restrictions, for instance, a Shopping Agent may only be able to perform one task at a time, may have limited time or an energy or fuel budget to perform a task, and so on. In \eu, these restrictions are represented by Constraints, which can be defined over any combination of Variables in a domain. \eu provides a Constraint Library that already implements many useful constraints (temporal, resource, relational, etc), it is also possible to add new constraints if required by a particular domain.

\item[\textbf{Objects}] The items we wish to describe and refer to in
  a domain are considered Objects. As in the case with object-oriented
  analysis and design, one can seek out the nouns in any domain
  description to find likely objects. In the Shopping Agent example,
  we might consider the Agent itself, Products and Locations all to be
  objects. Objects have state and behavior. For example, a Shopping
  Agent can have:

@@

  \item State: its location, a bag that contains the products it has
    already purchased (the bag can in turn be another object), etc.

@@

  Objects that have similar state and behavior can be
  described generically in terms of object types or classes. \eu
  allows the definition of classes in the same way that it is
  done in popular object-oriented programming languages, where a class 
  can have attributes (represented by Variables), and classes can be arranged in a single-parent hierarchy through inheritance. 
  However, one key aspect needs to be introduced to class definition in \eu that is not found 
  in object-oriented programming:  to describe state and behavior for the purposes of planning we need to
  build on the formalism of first order logic as explained below.

@@

\item[\textbf{Rules}] In order for a plan to be valid, it must comply with all rules and
regulations pertinent to the application domain in question. Rules
govern the internal and external relationships of a token. For
example, consider a parameterized action \texttt{Go(origin,destination)} which describes
a Shopping Agent traveling from one location ( \textit{origin}) to another ( \textit{destination}). The
parameters are instantiated on a token as variables whose domain of
values is the set of all locations in a given problem. A rule
governing an internal relationship among token variables might
stipulate that a transition must involve a change in location. This
can be easily expressed as a constraint on the definition of a
predicate of the form: \textit{origin != destination}. It makes sense to further
stipulate that the agent must be located at  \textit{origin} before travel can happen
and the agent must be located at  \textit{destination} when completed. This is
an example of a rule governing an external relationship among
tokens. It specifies a requirement that tokens of the predicate
\texttt{At} that represent the location state for the Shopping Agent 
precede and succeed tokens of the predicate \texttt{Go} that represents traveling.

\comment{ Add figure showing internal and external relationships from rules on action Go}

Figure xx illustrates the entities and relations involved in specifying
such a rule on a \texttt{Go} action. The token on which the rule applies
is referred to (\texttt{Go}) as the master. Each  \texttt{At} token required by the
master is referred to as a slave. All variables are indicated by name
and their domains are expressed as intervals in the case of temporal
variables and as enumerations for the remainder. Application of a rule
on a token can thus cause slave tokens, variables, and constraints to
be introduced.

The capability of a domain rule to cause a slave token to be created
is a key vehicle through which planning occurs. Semantically, this
rule imposes a requirement for supporting (slave) tokens to be in the plan in
order for the master to be valid.

\comment {talk about condition/effect annotations?}

@@

We have just covered the main elements of \eu's planning representation:

\begin{enumerate}
   \item \textbf{Domains, Variables and Objects} to represent the entities (Objects) and attributes (Variables) needed to represent a problem domain	
    \item \textbf{Tokens} Temporally scoped predicates to represent states and actions in time. 
    \item \textbf{Token States} which are necessary for the implementation of planning operations on a partial plan.
    \item \textbf{Constraints} to describe relationships among attributes of objects and tokens. \eu has a built-in library of constraints that covers time, resources and others that are common in real life problem domains.
    \item \textbf{Domain Rules} to describe internal and external relationships on and
     between tokens respectively. Rules are applied to active tokens (master tokens) 
     referred to as masters and typically produce inactive tokens (slaves) which are a key vehicle through which the planning
     process evolves. \comment{citation, or perhaps remove this statement?:}The rule structure in \eu differs from the more restrictive commitment to preconditions and effects in classical
     planning which prohibits durative actions and disembodied effects (i.e. effects which may not occur until some temporal distance after
     the end of the action).    
    \item \textbf{Timelines and Resources} are built in implementations of Object types that appear frequently in real life domains.
\end{enumerate}

%\begin{figure*}
%\centering 
%\subfloat[]{\label{fig:plan-evolve1}\includegraphics[width=0.3\textwidth]{figs/Plan-evolve-1.pdf}} 
%\subfloat[]{\label{fig:plan-evolve2}\includegraphics[width=0.3\textwidth]{figs/Plan-evolve-2.pdf}} 
%\subfloat[]{\label{fig:plan-evolve3}\includegraphics[width=0.3\textwidth]{figs/Plan-evolve-3.pdf}} 
%\caption{\small An illustrative plan synthesis example with concurrent
%  timelines. Each timeline is an instantiation of a subsystem tracked
%  by the planner. Tokens describe the instantiated state of subsystem
%  at a particular time instance and enforce temporal and parametric
%  constraints. Abstract goals are decomoposed into successively less
%  abstract tokens and instantiated in co-temporal timelines using
%  \texttt{Allen Algebra} relations. All tokens represent flexible
%  start/end times. \ref{fig:plan-evolve1} shows an initial state
%  evolving into \ref{fig:plan-evolve2} and \ref{fig:plan-evolve3}.}
%  \label{fig:Plan-evolve}
%  \vskip-5pt
%\end{figure*}

%Fig. \ref{fig:Plan-evolve} shows an illustrative example. We show four
%timelines which track \texttt{Goal, Path, Navigation} and
%\texttt{Command} state over time. Tokens on the goal timeline indicate
%a survey within a bounded volume and a min/max depth envelope. As the
%Volume Survey sub-goals on the \texttt{Path} timeline, the initial
%token on that timeline gets \emph{squeezed} to have dependencies of
%the goal token instantiated. Initially this dependency are the tokens
%\texttt{Go(x1,y1)} and \texttt{Go(xn,yn)} indicative of the area of
%coverage. The first of these in turn generates a sub-goal on the same
%timeline to go to the next intermediate waypoint, which is illustrated
%in Fig. \ref{fig:plan-evolve2}. Subsequent sub-goals (not shown in the
%figure) ultimately generate all the tokens to completion for this
%timeline. Fig. \ref {fig:plan-evolve3} shows a snapshot further along
%in plan generation and shows additional subgoals on the \texttt{Path,
%  Navigation} and \texttt{Command} timelines. Each token can thus
%trace its causality when instantiated in the plan. The domain model
%provides source of these temporal and parametric dependencies in the
%form of rules.

%\begin{minipage}[c]{\textwidth}
%\vspace{+0.5cm}
% \framebox[\textwidth][t]{
%Volume\_Survey(x$_1$,y$_1$,x$_n$,y$_n$,Min\_Depth, Max\_Depth)\\
%$\Rightarrow$ \\
%met\_by Go(x$_1$,y$_1$,Min\_Depth); \\
%starts [50,0] Go(x$_2$,y$_2$,Min\_Depth,Max\_Depth); \\
%\ldots{} \\
%ends [50,0] Go(x$_{n-1}$,y$_{n-1}$, Min\_Depth,Max\_Depth); \\
%meets Go(x$_n$,y$_n$,Min\_Depth, Max\_Depth); 
%\end{minipage}
